---
title: "Correlation"
output: html_notebook
---

## Pearson's correlation
$$r=\frac{\sum_{n}^{i=1}(x_i-\bar{x})(y_i-\bar{y})}{(n-1)s_{x}s_{y}}$$


### Problem
* Anxiety and exam performance
* Participants:
    + 103 students
* Measures
    + Exam performance (%)
    + Exam Anxiety (the EAQ, score out of 100)
    + Time spent revising (hours)


### Solution
```{r}
url <- "data/ExamAnxiety.csv"
examData <- read.csv(url,header = TRUE)
library(knitr)
## display first 10 rows
kable(head(examData,n = 10))
plot(examData$Anxiety,examData$Exam,xlab = "Anxiety",ylab = "Exam Score")
#abline(lm(examData$Exam~examData$Anxiety))
cor(examData,use = "complete.obs",method = "pearson")
cor.test(examData$Exam,examData$Anxiety,method = "pearson")
```

`alternative` indicates the alternative hypothesis and must be one of "two.sided", "greater" or "less". You can specify just the initial letter. "greater" corresponds to positive association, "less" to negative association.

```{r}
cor.test(examData$Exam, examData$Anxiety, alternative = "less", method = "pearson")
```

## Significance test 

To test the null hypothesis that the true correlation coefficient, $\rho$, is equal to zero, the data for each variable are assumed to come from normal distributions.

$$t=\frac{r\sqrt{n-2}}{\sqrt{1-r^2}} $$

If the null hypothesis is true, this statistic has a t-distribution, with n – 2 degrees of freedom.


```{r}
result <- cor.test(examData$Exam, examData$Anxiety, alternative = "less", method = "pearson")
str(result)
r <- result$estimate
n <- nrow(examData)
t <- r*sqrt(n-2)/sqrt(1-r^2)
t
pt(t,df = n-2)
curve(dt(x,df = n-2),from = -6,to = 6)
t.critical <- qt(0.025,df = n-2)
t.critical
points(t.critical,dt(t.critical,df = n-2),pch=16,col="red")
points(t,dt(t,df = n-2),pch=16,col="blue")
```

## Correlation and sample size
An extremely important point is that the correlation coefficient is influenced by sample size. It is far easier to reject the null hypothesis that $\rho=0$ with a large sample size than it is with a small sample size.

For large sample size, critial t value is approximately $2/\sqrt{n}$

```{r}
n <- c(10,15,20,30,50,100,250,500,1000)
r <- round(2/sqrt(n),3)
df <- data.frame(n,r)
names(df) <- c("Sample size", "Minimum absolute value of r needed to attain significance")
library(knitr)
kable(df)
```


## Spearman's correlation
* Spearman’s rho
    + Pearson’s correlation on the ranked data

$$r_{S} = 1 - \frac{6\sum_{i=1}^{n}d_{i}^{2}}{n^3-n}$$

### Problem
World’s Biggest Liar competition

* 68 contestants
* Measures
    + Where they were placed in the competition (first, second, third, etc.)
    + Creativity questionnaire (maximum score 60)


### Solution
```{r}
url <- "data/TheBiggestLiar.csv"
liarData <- read.csv(url,header = TRUE)
library(knitr)
## display first ten rows
kable(head(liarData,n = 10))
plot(liarData$Creativity,liarData$Position)
cor(liarData$Position, liarData$Creativity, method = "spearman")
cor.test(liarData$Position, liarData$Creativity, alternative = "less", method = "spearman")

```
