---
title: "Regression"
output: html_notebook
---

## Example 1

### Problem 

A record company boss was interested in predicting record sales from advertising.

* Data:

    + 200 different album releases
  
* Outcome variable:

    + Sales (CDs and downloads) in the week after release

* Predictor variable:

    + The amount (in units of $1000) spent promoting the record before release.


### Solution

```{r}
url <- "data/AlbumSales1.csv"
df <- read.csv(url,header = TRUE)
library(knitr)
kable(head(df,n = 10))

plot(df)
m <- lm(df$sales ~ df$adverts)
m
m <- lm(sales ~ adverts, data = df)
summary(m)
abline(m,col="red",lwd=2)
attributes(m)

```

The equation is: $sales = a + b\times adverts = 134.1399 + 0.0961 \times adverts$

## Example 2

Speed and Stopping Distances of Cars

```{r}
df <- cars
library(knitr)
kable(head(df,n = 10))
plot(df)
x <- cars$speed
y <- cars$dist
plot(y ~ x)
m <- lm(y~x)
m
summary(m)

abline(m,col="red",lwd=2)
segments(x,fitted(m),x,y,col="blue",lty = "dashed")

result <- cor.test(x,y,method = "pearson")
result
r <- result$estimate
r2 <- r^2
round(r2,4)
p=1
n=nrow(df)
adjusted.r2 <- r2-(1-r2)*p/(n-p-1)
round(adjusted.r2,4)

```


Make predictions based on new inputs:

```{r}
df <- cars
m <- lm(dist~speed,data = df)
n <- nrow(df)
df$x <- sample(10:100,size = n)
df$y <- predict(m,newdata = data.frame(speed=df$x))
library(knitr)
kable(head(df,n = 10))
plot(df$x,df$y)
abline(m,col="red")

```


## Example 3


```{r}
x <- c(65,35,30,44,80,77,32,39,44,77)
y <- c(120,68,35,60,100,91,44,71,89,113)

library(knitr)
kable(cbind(y,x))

n <- length(x)
mx <- mean(x)
my <- mean(y)
sx <- sd(x)
sy <- sd(y)

numerator <- sum((x-mx)*(y-my))
numerator
denominator <- sum((x-mx)^2)
denominator

slope <- numerator / denominator
slope
intercept <- my - slope * mx
intercept

m <- lm(y~x)
summary(m)

MSS <- sum((fitted(m) - my)^2)
MSS
RSS <- sum((y - fitted(m))^2)
RSS
TSS <- RSS + MSS
TSS

df1 <- 1
df2 <- n-2 
df3 <- df1+df2

msq1 <- MSS / df1
msq2 <- RSS / df2
F.value <- msq1 / msq2 
F.value
F.critical <- qf(0.95,df1 = 1,df2 = n-2)
F.critical
rsq <- MSS / TSS
rsq
if(F.value>F.critical){
  print("We reject the null hypothesis")
} else {
  print("We cannot reject the null hypothesis")
}

```

|          | Sum of squares |  df   | Mean square     |   F                |
|----------|----------------|-------|-----------------|--------------------|
|Regression|`r round(MSS,2)`|`r df1`|`r round(msq1,2)`|`r round(F.value,2)`|
|Residual  |`r round(RSS,2)`|`r df2`|`r round(msq2,2)`|                    |
|Total     |`r round(TSS,2)`|`r df3`|                 |                    |

